Tutorial_2a

### 2024-11-26 Tutorial_2a.Rmd
add figures of PPC for 2 models

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data. The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent). We perform some prior predictive checks to infer the prior distributions we will employ. We compare models using WAIC and LOO, and interpret parameter estimates for one model. 
We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.

@@

Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflects our prior beliefs. First, sample the prior distributions using sample_prior="only".

@@

```{r ppc-manual}
# prior predictive check for 1 time bin
# see http://bruno.nicenboim.me/bayescogsci/

@@

The shape of the prior predictive distribution looks as expected (compare with Figure X in the supplementary material): hazard values below .5 are more likely than hazard values above .5.

@@

@@

```{r inspect-M0i}
model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
summary(model_M0i)
formula(model_M0i)
plot(model_M0i)
```

@@

Perform a prior predictive check to see if the binary observations generated from the sprecified prior distributions reflect our prior beliefs. 
The following code allows to change the coefficients and generate (cloglog-)hazard functions.
Once you have found the coefficients for period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to estimate a mean and standard deviation for each coefficient.

@@

First, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random variables while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.

@@

```{r inspect-M0c-prior}
model_M0c_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior.rds")
```

@@

Generate prior predictive check for each participant.

@@

@@

Now increase the variance again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.

>>>>

Tutorial 2a shows various ways to fit Bayesian binary regression models to discrete time-to-event data. The data are taken from the first experiment of Panis and Schmidt (2016), but using only the no-mask trials (with factor prime = blank, congruent, or incongruent). We perform some prior predictive checks to determine the prior distributions we will employ. We compare models using WAIC and LOO, and interpret parameter estimates for one model. 
We end by plotting the logit and cloglog link functions, and illustrate how various prior distributions on the logit and cloglog scales look on the original probability (i.e., hazard) scale.

@@

Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflects our prior beliefs. First, sample the prior distributions using sample_prior="only".

@@

```{r ppc-manual, eval=T}
# prior predictive check for 1 time bin
# see http://bruno.nicenboim.me/bayescogsci/

@@

```{r save-ppc, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M0i_ppc_1bin.png", width = 10, height = 8, dpi = 800)
```

The shape of the prior predictive distribution looks as expected (compare with Figure 16 (row F) in the supplementary material): hazard values below .5 are more likely than hazard values above .5.

@@

```{r save-ppc-ten-bins-six-subjects, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M0i_ppc.png", width = 10, height = 8, dpi = 800)
```

@@

```{r inspect-M0i}
model_M0i <- readRDS("Tutorial_2_Bayesian/models/model_M0i.rds")
summary(model_M0i)
plot(model_M0i)
```

@@

Perform a prior predictive check to see if the binary observations generated from the specified prior distributions reflect our prior beliefs. 
The following code allows to change the coefficients and generate (cloglog-)hazard functions.
Once you have found the coefficients for the intercept, period_9 and period_9_sq that generate a variety of hazard functions spanning your prior beliefs, you can then use them to determine a suited mean and standard deviation for each coefficient.

@@

Next, sample the prior distributions, and set the variance of the normal prior for class "sd" to a low value, to minimize the effect of the random effects while inspecting the relation between the selected priors for period_9 and period_9_sq and the shape of the prior predictive distributions.

@@

```{r load-M0c-prior}
model_M0c_prior <- readRDS("Tutorial_2_Bayesian/models/model_M0c_prior.rds")
```

@@

Next, generate prior predictive check for each participant.

@@

```{r save-ppc-M1c, eval=F}
ggsave("Tutorial_2_Bayesian/figures/M0c_ppc.png", width = 10, height = 8, dpi = 800)
```

@@

Now increase the standard deviation again of the normal prior for class "sd", and fit model M0c to get the posterior distributions.
